---
title: "Data_Merge"
author: "Fiona R Lodge"
date: "May 10, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Statement

The purpose of this project is to develop insight into weather patterns that may trigger migraines.  Since barometric pressure is often a trigger, this variable was studying with more intent than the other meteorlogical data.  Data was collected from the NOAA by submitting an order at the site: https://www.ncdc.noaa.gov/cdo-web/datasets.  The sample chosen was 'local climatology data'.  Data was chosen from 20 cities in the US from 2013-2018.  The output of the order was in a csv format, and R was used as the chosen program.

## Data Wrangling to merge csvs
The first step of the data wrangling process was to merge the csvs from the 20 cities into one csv.  First, the csvs were moved into one folder called 'States_Data', and the working directory was set.

```{r setwd}
setwd("~/Documents/Masters_Year_2/weather_project_final/States_Data")
```

Then, all filenames were read into a list from the directory.

```{r list}
filenames <- list.files(full.names = TRUE)
```


A function was created to read in each file as a csv.
```{r readcsvs}
All <- lapply(filenames, function(i){
  read.csv(i, header = FALSE, skip = 1)})
```

All the csvs were merged on rows.
```{r rbind}
df_ALL <- do.call(rbind.data.frame, All)
```

I had to retype all of the names of the header of the csv.

```{r names}
names(df_ALL) <- c('STATION', 'STATION_NAME',	'ELEVATION',	
'LATITUDE', 'LONGITUDE',	'DATE',	'REPORTTPYE',	'HOURLYSKYCONDITIONS',
'HOURLYVISIBILITY',	'HOURLYPRSENTWEATHERTYPE',	'HOURLYDRYBULBTEMPF',
'HOURLYDRYBULBTEMPC',	'HOURLYWETBULBTEMPF',	'HOURLYWETBULBTEMPC',
'HOURLYDewPointTempF',	'HOURLYDewPointTempC',	'HOURLYRelativeHumidity',
'HOURLYWindSpeed',	'HOURLYWindDirection',	'HOURLYWindGustSpeed',
'HOURLYStationPressure',	'HOURLYPressureTendency',	'HOURLYPressureChange',
'HOURLYSeaLevelPressure',	'HOURLYPrecip',	'HOURLYAltimeterSetting',
'DAILYMaximumDryBulbTemp',	'DAILYMinimumDryBulbTemp',	'DAILYAverageDryBulbTemp',
'DAILYDeptFromNormalAverageTemp',	'DAILYAverageRelativeHumidity','DAILYAverageDewPointTemp',
'DAILYAverageWetBulbTemp',	'DAILYHeatingDegreeDays',	'DAILYCoolingDegreeDays',	'
DAILYSunrise',	'DAILYSunset',	'DAILYWeather',	'DAILYPrecip',	'DAILYSnowfall',
'DAILYSnowDepth',	'DAILYAverageStationPressure',	'DAILYAverageSeaLevelPressure',
'DAILYAverageWindSpeed',	'DAILYPeakWindSpeed',	'PeakWindDirection',
'DAILYSustainedWindSpeed',	'DAILYSustainedWindDirection',	'MonthlyMaximumTemp',
'MonthlyMinimumTemp',	'MonthlyMeanTemp',	'MonthlyAverageRH',	'MonthlyDewpointTemp',
'MonthlyWetBulbTemp',	'MonthlyAvgHeatingDegreeDays',	'MonthlyAvgCoolingDegreeDays',
'MonthlyStationPressure',	'MonthlySeaLevelPressure','MonthlyAverageWindSpeed',
'MonthlyTotalSnowfall',	'MonthlyDeptFromNormalMaximumTemp',	'MonthlyDeptFromNormalMinimumTemp',
'MonthlyDeptFromNormalAverageTemp',	'MonthlyDeptFromNormalPrecip',
'MonthlyTotalLiquidPrecip',	'MonthlyGreatestPrecip',	'MonthlyGreatestPrecipDate',
'MonthlyGreatestSnowfall',	'MonthlyGreatestSnowfallDate',	'MonthlyGreatestSnowDepth',
'MonthlyGreatestSnowDepthDate',	'MonthlyDaysWithGT90Temp',	'MonthlyDaysWithLT32Temp',
'MonthlyDaysWithGT32Temp',	'MonthlyDaysWithLT0Temp',	'MonthlyDaysWithGT001Precip',
'MonthlyDaysWithGT010Precip',	'MonthlyDaysWithGT1Snow',	'MonthlyMaxSeaLevelPressureValue',
'MonthlyMaxSeaLevelPressureDate',	'MonthlyMaxSeaLevelPressureTime',	
'MonthlyMinSeaLevelPressureValue',	'MonthlyMinSeaLevelPressureDate',
'MonthlyMinSeaLevelPressureTime',	'MonthlyTotalHeatingDegreeDays',	
'MonthlyTotalCoolingDegreeDays',	'MonthlyDeptFromNormalHeatingDD',	
'MonthlyDeptFromNormalCoolingDD',	'MonthlyTotalSeasonToDateHeatingDD',	
'MonthlyTotalSeasonToDateCoolingDD')
```

The next step was a bit of a moot point as I had to redo it in python, but changed the date format in place.

```{r date}
df_ALL['DATE'] = as.POSIXct(df_ALL$DATE)
```

View the dim and str of the data frame to make sure it makes sense.
```{r dim}
dim(df_ALL)
str(df_ALL)
```

Write the dataframe as a csv. Note: this had already been done, so weather_data.csv is the file used elsewhere.
```{r csv}
#write.csv(df_ALL, "weather_data.csv")
```

